Centers and centralizers of higher enveloping algebras of Lie algebras

### 2020-09-09 draft.tex
harish chandra

\begin{proof}
A cochain complex $V^\bullet$ is an abelian dg Lie algebra;
indeed, so is any shift $V^\bullet[k]$. 
Let $\fg_V$ denote $V^\bullet[-n+1]$ viewed as an abelian dg Lie algebra.
The Poincar\'e lemma for compactly supported de Rham forms on $\RR^n$ provides a canonical quasi-isomorphism
\[
\Omega^*_c(\RR^n) \xto{\simeq} \RR[-n]
\]
by the natural quotient map and hence determines a canonical quasi-isomorphism
\[
\cliels(\Omega^*_c(\RR^n) \otimes \fg_V) \xto{\simeq} \Sym(V^\bullet).
\]
This map makes sense on any ball in $\RR^n$ and thus implies that $\bU_n \fg_V$ is quasi-isomorphic as a $\Disk_n$ algebra to~$\Sym(V^\bullet)$.
Applying our central theorem, we see
\begin{align*}
\bZ_n(\fg_V) &= \clies(\Omega^* \otimes \fg_V, \bU_n \fg_V) \\
&= \Sym((\Omega^* \otimes \fg_V)^\vee[-1]) \otimes \Sym(\Omega^*_c \otimes \fg_V[1])\\
& \simeq \Sym(V^\bullet \oplus (V\bullet)^\vee[-n])
\end{align*}
because
\[
\Omega^*(\RR^n) \otimes \fg_V \simeq V[-n+1]
\]
so
\[
\Sym( (\Omega^*(\RR^n) \otimes \fg_V)^\vee[-1]) \simeq \Sym( V^\vee[-n]).
\]
This identification concludes the argument.
\end{proof}

@@

\begin{prp}
For $\fg$ a reductive \owen{semisimple?} Lie algebra with Cartan subalgebra~$\fh$ and Weyl group~$W$, 
there is a quasi-isomorphism of $E_{n+1}$ algebra
\[
{\frak Z}_n(\UU_n \fg) \simeq H^*_{\rm Lie}(\fg) \otimes \Sym(\fh[1-n])^W.
\]
\end{prp}

\begin{proof}
\owen{In brief
\[
{\frak Z}_n(\UU_n \fg) \simeq \clies(\fg, \Sym(\fg[1-n])
\]
and then apply theorems about Lie algebra cohomology plus Chevalley restriction theorem.}
\end{proof}

>>>>

\begin{proof}
A cochain complex $V^\bullet$ is an abelian dg Lie algebra;
indeed, so is any shift $V^\bullet[k]$. 
Let $\fg_V$ denote $V^\bullet[-n+1]$ viewed as an abelian dg Lie algebra.
The Poincar\'e lemma for compactly supported de Rham forms on $\RR^n$ provides a canonical quasi-isomorphism
\[
\Omega^\bu_c(\RR^n) \xto{\simeq} \RR[-n]
\]
by the natural quotient map and hence determines a canonical quasi-isomorphism
\[
\cliels(\Omega^\bu_c(\RR^n) \otimes \fg_V) \xto{\simeq} \Sym(V^\bullet).
\]
This map makes sense on any ball in $\RR^n$ and thus implies that $\bU_n \fg_V$ is quasi-isomorphic as a $\Disk_n$ algebra to~$\Sym(V^\bullet)$.
Applying our central theorem, we see
\begin{align*}
\bZ_n(\fg_V) &= \clies(\Omega^* \otimes \fg_V, \bU_n \fg_V) \\
&= \Sym((\Omega^\bu \otimes \fg_V)^\vee[-1]) \otimes \Sym(\Omega^\bu_c \otimes \fg_V[1])\\
& \simeq \Sym(V^\bullet \oplus (V\bullet)^\vee[-n])
\end{align*}
because
\[
\Omega^\bu (\RR^n) \otimes \fg_V \simeq V[-n+1]
\]
so
\[
\Sym( (\Omega^\bu(\RR^n) \otimes \fg_V)^\vee[-1]) \simeq \Sym( V^\vee[-n]).
\]
This identification concludes the argument.
\end{proof}

@@

For a reductive Lie algebra $\fg$, the Harish-Chandra isomorphism is an isomorphism of commutative algebras
\[
Z(U(\fg)) \xto{\cong} {\rm Sym} (\fh)^W
\]
where $\fh$ is a Cartan subalgebra and $W$ is the Weyl group of $\fg$. 
Here $Z(U(\fg))$ denotes the ordinary center of the associative algebra $U(\fg)$.

The following result generalizes (and implies) the Harish--Chandra isomorphism for the $\EE_n$-center of the $\EE_n$-enveloping algebra.

\begin{prp}
For $\fg$ a reductive \owen{semisimple?} Lie algebra with Cartan subalgebra~$\fh$ and Weyl group~$W$, 
there is a quasi-isomorphism of $\EE_{n+1}$-algebras
\[
{\frak Z}_n(\UU_n \fg) \simeq H^\bu_{\rm Lie}(\fg) \otimes \Sym(\fh[1-n])^W.
\]
\end{prp}

\brian{Is there an explicit description of the $\EE_{n+1}$-structure for the RHS?}

\begin{proof}
By the Poincar\'{e} lemma for de Rham forms and compactly supported de Rham forms, one has a canonical quasi-isomorphism
\[
\bZ_n(\UU_n \fg) \xto{\simeq} \clies\left(\fg, \Sym(\fg[1-n]) \right) .
\]
Since $\fg$ is reductive, there is a quasi-isomorphism \brian{do we actually need semisimple here?}
\[
\clies(\fg, \Sym(\fg[1-n]) \simeq H^\bu_{\rm Lie} (\fg) \otimes \Sym(\fg[1-n])^G 
\]
where $G$ is a compact connected Lie group whose Lie algebra is $\fg$. 
Further, by the Chevalley restriction theorem one has 
\[
\Sym(\fg[1-n])^G \cong \Sym \left(\fh[1-n]\right)^W .
\]
\end{proof}
